Behavior Segmentation for Smartphone Use 

### 2025-01-22 ProjectBlog/paper1/Clustering Anlysis/cluster.qmd
interflex
cluster analysis

# Creating a special bin for 0
df1=df[df['screen_time']>15]
bins = [-np.inf, 0.001] + list(range(15, 185, 5))  # 0.1 to include 0 in the first bin exclusively
df1['screen_time_bin'] = pd.cut(df1['screen_time'], bins=bins, right=False, include_lowest=True)
# Recalculating the aggregated data
grouped = df1.groupby('screen_time_bin').agg({'sleep_debt':'mean', 'message':'mean', 'screen_time':'count'}).reset_index()
# Replotting with the adjusted bins
fig, ax1 = plt.subplots(figsize=(14, 8))
# Line plot for sleep debt trend
line1, = ax1.plot(grouped['screen_time_bin'].astype(str), grouped['sleep_debt'], color='blue', linestyle='-', marker='o', label='Sleep Debt')
# Adding labels for sleep debt
for x, y in zip(grouped['screen_time_bin'].astype(str), grouped['sleep_debt']):
    ax1.text(x, y, f'{y:.2f}', va='bottom', ha='center', color='blue')
# Setting the left y-axis for sleep debt
ax1.set_xlabel('Screen Time (minutes)')
ax1.set_ylabel('Average Sleep Debt')
ax1.set_title('Sleep Debt, Screen Time and Message')
ax1.tick_params(axis='x', rotation=45)
# Line plot for message count on a secondary y-axis
ax2 = ax1.twinx()
line2, = ax2.plot(grouped['screen_time_bin'].astype(str), grouped['message'], color='darkorange', linestyle='--', marker='o', label='Message')
# Adding labels for message count
for i, txt in enumerate(grouped['message']):
    ax2.annotate(f'{txt:.2f}', (grouped['screen_time_bin'].astype(str)[i], grouped['message'][i]), textcoords="offset points", xytext=(0,10), ha='center')
ax2.set_ylabel('Message')
# Bar plot for the number of observations on a secondary y-axis
ax3 = ax1.twinx()
bars = ax3.bar(grouped['screen_time_bin'].astype(str), grouped['screen_time'], width=0.4, color='lightgrey', alpha=0.5, label='Number of Observations')
ax3.spines['right'].set_position(('outward', 60))  # Offset the right spine of ax3
ax3.set_ylabel('Number of Observations')
# Labeling the number of observations above each bar
for bar in bars:
    yval = bar.get_height()
    ax3.text(bar.get_x() + bar.get_width() / 2, yval, f'{int(yval)}', va='bottom', ha='center', color='black')
# Adjusting the plot
ax1.grid(True, which='both', linestyle='--', linewidth=0.5)
ax2.grid(False)
ax3.grid(False)
# Adding legends
fig.legend(loc='upper right')
fig.tight_layout()
plt.show()
```

>>>>

import pandas as pd
import matplotlib.pyplot as plt
from matplotlib.lines import Line2D

# Assuming df1 is already defined as df
df1 = df

# Data preparation steps (as per your instructions)
df1['screen_time_bin'] = None

# Creating bins as [0, 10), [10, 20), ...
bins = list(range(0, 190, 10))  # Define bins from 0 to 180 in steps of 10
labels = [f"[{bins[i]},{bins[i+1]})" for i in range(len(bins)-1)]  # Create labels like '[0,10)', '[10,20)', etc.

# Creating screen_time_bin column using pd.cut
df1['screen_time_bin'] = pd.cut(df1['screen_time'], bins=bins, labels=labels, right=False, include_lowest=True)

# Aggregating data for each bin
grouped = df1.groupby('screen_time_bin').agg({
    'sleep_debt': 'mean',
    'message': 'mean',
    'screen_time': 'count'
}).reset_index()

# Plotting
plt.rcParams['font.family'] = 'Times New Roman'
fig, ax1 = plt.subplots(figsize=(14, 8))

# Line plot for sleep debt trend
line1, = ax1.plot(grouped['screen_time_bin'], grouped['sleep_debt'], color='purple',
                  linestyle='-', marker='o', label='Sleep Debt')
for x, y in zip(grouped['screen_time_bin'], grouped['sleep_debt']):
    ax1.annotate(f'{y:.2f}', (x, y), textcoords="offset points", xytext=(0, 10), ha='center', color='purple')

# X and Y labels
ax1.set_xlabel('Screen Time (minutes)', fontsize=14, color='black')
ax1.set_ylabel('Sleep Debt (minutes)', fontsize=14, color='black')
ax1.tick_params(axis='x', rotation=30, labelcolor='black', labelsize=10)
ax1.tick_params(axis='y', labelcolor='black', labelsize=12)
# ax1.set_title('Sleep Debt, Screen Time and Message')

# Secondary axis for message count
ax2 = ax1.twinx()
line2, = ax2.plot(grouped['screen_time_bin'], grouped['message'], color='darkorange',
                  linestyle='-', marker='o', label='Message')
for i, txt in enumerate(grouped['message']):
    ax2.annotate(f'{txt:.2f}', (grouped['screen_time_bin'][i], grouped['message'][i]),
                 textcoords="offset points", xytext=(0, 10), ha='center', color='darkorange')
ax2.set_ylabel('Message', fontsize=14, color='black')
ax2.tick_params(axis='y', labelcolor='black', labelsize=12)

# Tertiary axis for bar plot of observations
ax3 = ax1.twinx()
bars = ax3.bar(grouped['screen_time_bin'], grouped['screen_time'], width=0.4, color='lightblue',
               alpha=0.3, label='Number of Observations')
ax3.spines['right'].set_position(('outward', 60))

# Hide the axis for Number of Observations
ax3.get_yaxis().set_visible(False)
ax3.spines['right'].set_visible(False)

# Draw connection lines from bar bottoms to the x-axis range
for bar, bin_label in zip(bars, grouped['screen_time_bin']):
    bar_x = bar.get_x() + bar.get_width() / 2
    bar_bottom = bar.get_y()
    ax1.add_line(Line2D([bar_x, bar_x], [bar_bottom, -10], color='gray', linestyle='-'))

for bar in bars:
    yval = bar.get_height()
    ax3.text(bar.get_x() + bar.get_width() / 2, yval, f'{int(yval)}', va='bottom', ha='center', color='black')

# Legend
lines = [line1, line2]
bars_proxy = plt.Line2D([0], [0], color='lightblue', alpha=0.3, lw=8)
labels = ['Sleep Debt', 'Message', 'Number of Observations']
fig.legend(lines + [bars_proxy], labels, loc='upper right', bbox_to_anchor=(0.96, 0.99), fontsize=12)

# Tight layout and save
fig.tight_layout()
plt.savefig('/Users/shaokangyang/Downloads/model-free.svg', format='svg', dpi=300)
plt.show()
```

### **Marginal effect estimates with multiplicative interaction models**

![](images/nonparametric%20rela.svg){fig-align="center"}
